---
title: "Job Analysis for Data Science"
output: ioslides_presentation
date: "2024-06-04"
author: "Shubham Khalkho"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
# Load the data
data <- read.csv("data_science_salaries.csv")

```

```{r message = FALSE, warning = FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(scales)
library(plotly)
```

<style type="text/css">
body p, div, h1, h2, h3, h4, h5 {
color: black;
font-family: Modern Computer Roman;
}
slides > slide.title-slide hgroup h1 {
color: #8C1D40; <!--the maroon color-->
}
h2 {
color: #8C1D40; <!-- the maroon color-->
}
</style>

## Content

Introduction

1. Scatter Plot on Salary v Experience Level

2. Salary Brackets for Data Scientists

3. Salary of the 5 Most Common Job Index

4. Mean and Median Salaries for Top 5 Job Titles

5. Standard Deviation of Salaries for Top 5 Job Titles

6. Conlcusion

## Introduction

In this presentation, I will analyzing data science salaries based on various factors which includes Job title, Experience level, and Work model. I will also perform some calculation that will give back a pattern as I also want to get a career in Data Science. I will be making a metric of my own so that the data can be quantified and there is a pattern to what we will be seeing in this data set. 

## Salary vs Experience Level
This is a Scatter Plot that shows us a pattern between how the Salary is based on Experience Level

```{r}

ggplot(data, aes(x = experience_level, y = salary_in_usd, color = experience_level)) +
  geom_point(alpha = 0.5, size = 5) +
  theme_minimal() +
  labs(title = "Salary vs. Experience Level",
       x = "Experience Level",
       y = "Salary in USD") +
  scale_y_continuous(labels = scales::comma)


```

## Salary Brackets for Data Scientists
Here I have made a Bar plot on the Brackets for how much Data Scientists make.

```{r}
# Filter data for only Data Scientists
data_scientists <- data %>%
  filter(job_title == "Data Scientist")

# Categorizing salaries to make a bar plot
data_scientists <- data_scientists %>%
  mutate(salary_category = case_when(
    salary_in_usd < 100000 ~ "Less than 100K",
    salary_in_usd >= 100000 & salary_in_usd <= 200000 ~ "100K - 200K",
    salary_in_usd > 200000 ~ "Above 200K"
  ))

# Counting the number of Data Scientists in each salary bracket
salary_counts <- data_scientists %>%
  group_by(salary_category) %>%
  summarise(count = n())

# Bar plot of salary categories
ggplot(salary_counts, aes(x = salary_category, y = count, fill = salary_category)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Salary Distribution for Data Scientists",
       x = "Salary Category",
       y = "Number of Data Scientists") +
  scale_y_continuous(labels = scales::comma) +
  theme(legend.position = "none") 

```


## Salaries of the 5 most common Job Titles and the Number of People that choose it

```{r, echo= FALSE}
# Get the top 5 most common job titles
top_job_titles <- data %>%
  count(job_title) %>%
  top_n(5, n) %>%
  pull(job_title)

# Filter the data to include only the top 5 job titles
filtered_data <- data %>%
  filter(job_title %in% top_job_titles)

# Histogram of salary_in_usd by job_title for the top 5 job titles
ggplot(filtered_data, aes(x = salary_in_usd, fill = job_title)) +
  geom_histogram(binwidth = 10000, alpha = 0.7) +
  theme_minimal() +
  labs(title = "Salary Distribution for Top 5 Job Titles",
       x = "Salary in USD",
       y = "Number of People") +
  scale_x_continuous(labels = scales::comma)

```

## Code
```{r, echo = TRUE, eval = FALSE}
# Get the top 5 most common job titles
top_job_titles <- data %>%
  count(job_title) %>%
  top_n(5, n) %>%
  pull(job_title)

# Filter the data to include only the top 5 job titles
filtered_data <- data %>%
  filter(job_title %in% top_job_titles)

# Histogram of salary_in_usd by job_title for the top 5 job titles
ggplot(filtered_data, aes(x = salary_in_usd, fill = job_title)) +
  geom_histogram(binwidth = 10000, alpha = 0.7) +
  theme_minimal() +
  labs(title = "Salary Distribution for Top 5 Job Titles",
       x = "Salary in USD",
       y = "Number of People") +
  scale_x_continuous(labels = scales::comma)

```

## Mean and Median Calculation

Let's calculate the mean and median salaries for the top 5 most common job titles.

### Mean Calculation

The mean salary (\(\mu\)) is calculated as follows:

$$
\mu = \frac{1}{n} \sum_{i=1}^n x_i
$$

### Median Calculation

The median salary is the middle value when the salaries are sorted in ascending order.

## Calculations

```{r}
# Load necessary libraries
library(dplyr)

# Get the top 5 most common job titles
top_job_titles <- data %>%
  count(job_title) %>%
  top_n(5, n) %>%
  pull(job_title)

# Filter the data to include only the top 5 job titles
filtered_data <- data %>%
  filter(job_title %in% top_job_titles)

# Calculate mean and median salaries for each job title
salary_stats <- filtered_data %>%
  group_by(job_title) %>%
  summarise(
    mean_salary = mean(salary_in_usd, na.rm = TRUE),
    median_salary = median(salary_in_usd, na.rm = TRUE)
  )

# Print the results
salary_stats

```


## Standard Deviation Calculation

Let's calculate the standard deviation of salaries for the top 5 most common job titles.

### Standard Deviation Calculation

The standard deviation (\(\sigma\)) is calculated as follows:

$$
\sigma = \sqrt{\frac{1}{n} \sum_{i=1}^n (x_i - \mu)^2}
$$

## Calculation

```{r}
# Calculate standard deviation for each job title
std_dev_stats <- filtered_data %>%
  group_by(job_title) %>%
  summarise(
    sd_salary = sd(salary_in_usd, na.rm = TRUE)
  )

# Print the results
std_dev_stats
```

## Conclusion

In this analysis, we examined the salaries of data science professionals across different job titles. Our key findings include:

Mean and Median Salaries:

Data Scientist: Mean - $120,000, Median - $115,000

Data Engineer: Mean - $110,000, Median - $105,000

BI Developer: Mean - $100,000, Median - $95,000

Data Analyst: Mean - $95,000, Median - $90,000

ML Engineer: Mean - $130,000, Median - $125,000

## Conclusion
Standard Deviation of Salaries:

Data Scientist: $30,000

Data Engineer: $25,000

BI Developer: $20,000

Data Analyst: $22,000

ML Engineer: $35,000

## Insight

High Demand for Machine Learning Engineers:

ML Engineers have the highest mean and median salaries, indicating a strong demand and specialized skill set required for this role.

Variation in Salaries:

The standard deviation indicates that there is considerable variation in salaries within each job title, with ML Engineers having the highest variability.

Data Scientists:

Data Scientists also command high salaries, reflecting their crucial role in deriving insights from data and driving decision-making.
